---
title: "Assignment 2: Coding Basics"
author: "Katie Tack"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: inline
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics, Part 1

1.  Generate a sequence of numbers from one to 30, increasing by threes. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r basics 1}
#1. 
seq(1, 30, 3) #starting with 1 ending in 30, increasing by 3
thirty_sequence <- seq(1, 30, 3) #naming the sequence
thirty_sequence #same result as line 35, can be done without doing line 35 first
#2. 
mean(thirty_sequence) #find mean, doesn't need to be done for finding mean > median
median(thirty_sequence) #find median
thirty_mean <- mean(thirty_sequence) #assign name to mean so it can be used later
thirty_median <- median(thirty_sequence) #assign name to median so it can be used later
#3. 
thirty_mean > thirty_median #ask R if the mean is greater than the median
```

## Basics, Part 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r basics 2}
student_names <- c('Amanda','Brian','Jeff','Ana','Debbie','Ethan') #character
test_scores <- c(79,42,95,51,88,32) #numeric
if_passed <- test_scores >= 50
df_student_info <- data.frame(StudentName = student_names, StudentScores = test_scores, Passed = if_passed)
colnames <- c('Student Names', 'Test Scores', 'Passed')
```

9.  QUESTION: How is this data frame different from a matrix?

> Answer: This data frame is stored in a table with multiple different data types (character, numerical, and logical) in many columns, whereas a matrix is two-dimensional, can only be one type of data, and is homogenous. 

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r basics 3}

```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer:
