---
title: "Assignment 10: Data Scraping"
author: "Katie Tack"
output: pdf_document
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on data scraping. 

## Directions
1. Rename this file `<FirstLast>_A10_DataScraping.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure your code is tidy; use line breaks to ensure your code fits in the knitted output.
5. Be sure to **answer the questions** in this assignment document.
6. When you have completed the assignment, **Knit** the text and code into a single PDF file.


## Set up 
1. Set up your session:

* Load the packages `tidyverse`, `rvest`, and any others you end up using.
* Check your working directory

```{r 1. Project Setup, message = FALSE, warning=FALSE}
#1 
library(tidyverse)
library(rvest)
library(here)
library(ggplot2)
library(lubridate)
library(purrr)

here()


```

2. We will be scraping data from the NC DEQs Local Water Supply Planning website, specifically the Durham's 2022 Municipal Local Water Supply Plan (LWSP): 
 * Navigate to https://www.ncwater.org/WUDC/app/LWSP/search.php
 * Scroll down and select the LWSP link next to Durham Municipality. 
 * Note the web address: <https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2022>
 
Indicate this website as the as the URL to be scraped. (In other words, read the contents into an `rvest` webpage object.)

```{r set.the.scraping.website, warning=FALSE}
#2 
theURL <- 'https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=03-32-010&year=2022'
theWebpageObject <- read_html(theURL)
theWebpageObject

```

3. The data we want to collect are listed below:

* From the "1. System Information" section:
 * Water system name
 * PWSID
 * Ownership
 
* From the "3. Water Supply Sources" section:
 * Maximum Day Use (MGD) - for each month

In the code chunk below scrape these values, assigning them to four separate variables.

>HINT: The first value should be "Durham", the second "03-32-010", the third "Municipality", and the last should be a vector of 12 numeric values (represented as strings)".

```{r scrape.the.data, warning=FALSE}
#3 
WaterSystemName <- theWebpageObject %>%
  html_nodes("div+ table tr:nth-child(1) td:nth-child(2)") %>%
  html_text()
WaterSystemName

PWSID <- theWebpageObject %>%
  html_nodes("td tr:nth-child(1) td:nth-child(5)") %>%
  html_text()
PWSID

Ownership <- theWebpageObject %>%
  html_nodes("div+ table tr:nth-child(2) td:nth-child(4)") %>%
  html_text()
Ownership

MaxDayUse <- theWebpageObject %>%
  html_nodes("th~ td+ td") %>%
  html_text()
MaxDayUse

```


4. Convert your scraped data into a dataframe. This dataframe should have a column for each of the 4 variables scraped and a row for the month corresponding to the withdrawal data. Also add a Date column that includes your month and year in data format. (Feel free to add a Year column too, if you wish.)

>TIP: Use `rep()` to repeat a value when creating a dataframe.

>NOTE: It's likely you won't be able to scrape the monthly widthrawal data in chronological order. You can overcome this by creating a month column manually assigning values in the order the data are scraped: "Jan", "May", "Sept", "Feb", etc... Or, you could scrape month values from the web page...

5. Create a line plot of the maximum daily withdrawals across the months for 2022

```{r create.a.dataframe.from.scraped.data, warning=FALSE}
#4 
month <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
           "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

DurhamWater <- data.frame("Month" = month,
                          "Year" = rep(2022,12))

DurhamWater <- DurhamWater %>%
  mutate(
    WaterSystemName = WaterSystemName,
    PWSID = PWSID,
    Ownership = Ownership,
    MaxDayUse = MaxDayUse
  ) 
DurhamWater$MaxDayUse <- as.numeric(DurhamWater$MaxDayUse)

#5 
ggplot(DurhamWater,aes(x=Month,y=MaxDayUse, group = 1)) + 
  geom_line() +
  labs(
    title = "2022 Maximum Daily Withdrawls in Durham Municipality",
    subtitle = "Created by Katie Tack",
    y="Withdrawal (mgd)",
    x="Month",
  ) +
  scale_x_discrete(labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                              "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
  scale_y_continuous(breaks = seq(0, ceiling(max(DurhamWater$MaxDayUse)), by = 2.5))
  
```

6. Note that the PWSID and the year appear in the web address for the page we scraped. Construct a function using your code above that can scrape data for any PWSID and year for which the NC DEQ has data. **Be sure to modify the code to reflect the year and site (pwsid) scraped**.

```{r construct.a.scraping.function, warning=FALSE}
#6. 
scraping <- function(the_pwsid, the_year){
  NCWaterSite <- read_html(paste0('https://www.ncwater.org/WUDC/app/LWSP/report.php?pwsid=', 
                                  the_pwsid, '&year=', the_year))
  
  WaterSystemName_tag <- 'div+ table tr:nth-child(1) td:nth-child(2)'
  PWSID_tag <- 'td tr:nth-child(1) td:nth-child(5)'
  Ownership_tag <- 'div+ table tr:nth-child(2) td:nth-child(4)'
  MaxDayUse_tag <- 'th~ td+ td'
  
  WaterSystemName <- NCWaterSite %>% html_nodes(WaterSystemName_tag) %>% html_text()
  PWSID <- NCWaterSite %>% html_nodes(PWSID_tag) %>% html_text()
  Ownership <- NCWaterSite %>% html_nodes(Ownership_tag) %>% html_text()
  MaxDayUse <- NCWaterSite %>% html_nodes(MaxDayUse_tag) %>% html_text()
  
  month <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
           "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
  DurhamWater <- data.frame("Month" = month,
                            "Year" = rep(the_year,12))
  DurhamWater <- DurhamWater %>%
    mutate(
      WaterSystemName = WaterSystemName,
      PWSID = PWSID,
      Ownership = Ownership,
      MaxDayUse = MaxDayUse
    )
  DurhamWater$MaxDayUse <- as.numeric(DurhamWater$MaxDayUse)
  return(DurhamWater)
}


```

7. Use the function above to extract and plot max daily withdrawals for Durham (PWSID='03-32-010') for each month in 2015

```{r fetch.and.plot.Durham.2015.data, warning=FALSE}
#7 
Durham2015 <- scraping('03-32-010',2015)

ggplot(Durham2015,aes(x=Month,y=MaxDayUse, group = 1,)) + 
  geom_line() +
  labs(
    title = "2015 Maximum Daily Withdrawls in Durham Municipality",
    subtitle = "Created by Katie Tack",
    y="Withdrawal (mgd)",
    x="Month",
  ) +
  scale_x_discrete(labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                              "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
  scale_y_continuous(breaks = seq(0, ceiling(max(Durham2015$MaxDayUse)), by = 2.5))

```

8. Use the function above to extract data for Asheville (PWSID = 01-11-010) in 2015. Combine this data with the Durham data collected above and create a plot that compares Asheville's to Durham's water withdrawals.

```{r fetch.and.plot.Asheville.2015.data, warning=FALSE}
#8 
Asheville2015 <- scraping('01-11-010',2015)

combined2015 <- rbind(Asheville2015, Durham2015)

combined2015$MaxDayUse <- as.numeric(combined2015$MaxDayUse)


ggplot(combined2015,aes(x=Month,y=MaxDayUse, color=PWSID, group = PWSID)) + 
  geom_line() +
  labs(
    title = "2015 Maximum Daily Withdrawls in Durham & Asheville",
    subtitle = "Created by Katie Tack",
    y="Withdrawal (mgd)",
    x="Month"
  ) +
  scale_x_discrete(labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                              "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
  scale_y_continuous(breaks = seq(0, ceiling(max(combined2015$MaxDayUse)), by = 2.5))

```


9. Use the code & function you created above to plot Asheville's max daily withdrawal by months for the years 2010 thru 2021.Add a smoothed line to the plot (method = 'loess').

> TIP: See Section 3.2 in the "10_Data_Scraping.Rmd" where we apply "map2()" to iteratively run a function over two inputs. Pipe the output of the map2() function to `bindrows()` to combine the dataframes into a single one. 

```{r fetch.and.plot.Asheville.multiyear.data, warning=FALSE}
#9 
AshevillePWSID <- "01-11-010"

GetAsheville2010to2021 <- function(pwsid, year) {
  scraping(the_pwsid, the_year)
}

AshevilleYears <- 2010:2021

Asheville2010to2021 <- map2_dfr(
  AshevillePWSID, AshevilleYears, 
  ~ scraping(.x, .y),
  .id = "Year"
)

AshevilleMGD <- Asheville2010to2021 %>%
  mutate(Year = case_when(
    Year == 1 ~ "2010",
    Year == 2 ~ "2011",
    Year == 3 ~ "2012",
    Year == 4 ~ "2013",
    Year == 5 ~ "2014",
    Year == 6 ~ "2015",
    Year == 7 ~ "2016",
    Year == 8 ~ "2017",
    Year == 9 ~ "2018",
    Year == 10 ~ "2019",
    Year == 11 ~ "2020",
    Year == 12 ~ "2021",
  ))

ggplot(AshevilleMGD,aes(x=Month,y=MaxDayUse, color=Year, group = Year)) + 
  geom_line() +
  labs(
    title = "Maximum Daily Withdrawls in Asheville: 2010-2021",
    subtitle = "Created by Katie Tack",
    y="Withdrawal (mgd)",
    x="Month"
  ) +
  scale_x_discrete(labels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun",
                              "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")) +
  scale_y_continuous(breaks = seq(0, ceiling(max(AshevilleMGD$MaxDayUse)), by = 1))

#used ChatGPT as I wanted to see other ways to scrape in multiple years
#I was not a huge fan of the way they did it as it didn't bring in the actual years into the Year column
#which meant I had to reassign the rows into the correct format
#I decided to keep it in though as another option for myself in the future

```

Question: Just by looking at the plot (i.e. not running statistics), does Asheville have a trend in water usage over time?
> Answer:
>
